---
title: "ANOMALY DETECTION"
author: "KIPLANGAT-DAVID"
date: "4/3/2022"
output: html_document
---

##RESEARCH QUESTION##

Carrefour Kenya and are currently undertaking a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax).This project is aimed at doing analysis on the dataset provided by carrefour and create insights on how to achieve highest sales.

##METRIC FOR SUCCESS##

Be able to detect and do away with anomalies in our dataset

##THE CONTEXT##

Carre Four is an International chain of retail supemarkets in the world, It was set up in Kenya in the year 2016 and has been performing well over the years.Carrefour ensures customer satisfaction and everyday convenience while offering unbeatable value for money with a vast array of more than 100,000 products, shoppers can purchase items for their every need, whether home electronics or fresh fruits from around the world, to locally produced items.
This project is aimed at creating insights from existing and current trends to develop marketing strategies that will enable the marketing team achieve higher sales.



##EXPERIMENTAL DESIGN##

 1. Loading libraries
 2. Load data
 2. Data cleaning
 3. Exploratory analysis
 4. Anomaly detection
 5. Conclusion
 6. Recommendation
 

##DATA RELEVANCE##

The data we are using for this  study was sourced from CarreFour database and is a reflection of current transactions

##1. LIBRARIES##

```{r}
#pkg <- c(‘tidyverse’,’tibbletime’,’anomalize’,’timetk’)
#install.packages('anomalize')
library(tidyverse)
library(tibbletime)
library(anomalize)
library(timetk)
library(ggplot2)
library(data.table)
```

##2. LOAD DATA##
```{r}
df <- read.csv('http://bit.ly/CarreFourSalesDataset')
```

```{r}
#preview dataset
head(df)
```

##Data structure##
```{r}
#data structure
str(df)
```
 - our dataset has 1000 observations and 2 variables
 - our date column is character type. We will transform to date type

```{r}
#convert date column to date type
df$Date <- as.Date(df$Date,'%m/%d/%y')
str(df)
```
Next,
- Convert date to #POCIXct# type
POSIXct stores both a date and time with an associated time zone. The default time zone selected, is the time zone that your computer is set to which is most often your local time zone
```{r}
df$Date <- as.POSIXct(df$Date)
```

-Check class
```{r}
#class
class(df)
```
We will convert our data to class tibble because for Anomaly Detection using anomalize, we need to have either a tibble or tibbletime object

```{r}
df <- as_tibble(df)
```

```{r}
df <- na.omit(df)
```

##MISSING VALUES##

```{r}
sum(is.na(df))
```
##DUPLICATES##

```{r}
anyDuplicated(df)
```
##OUTLIERS##

```{r}
boxplot(df$Sales)
```

- The data has outliers beyond maximum

##3. EXPLORATORY DATA ANALYSIS##

Plot sales against date

```{r}
ggplot(df, aes(x=Date, y=Sales, color=Sales)) + geom_line()
```

##4. ANOMALY DETECTION##

The R ‘anomalize’ package enables a workflow for detecting anomalies in data. The main functions are:
 - time_decompose()- for time series decomposition
 - anomalize()- for Anomaly detection of remainder
 - time_recompose()- for Anomaly lower and upper bound transformation

##Decomposing##

We will use time_decompose() function in anomalize package. We will use stl method which extracts seasonality

```{r}
df %>%
time_decompose(Sales, method = 'stl', frequency = 'auto', trend = 'auto') %>%
anomalize(remainder, method = 'gesd', alpha = 0.1, max_anoms = 0.5) %>%
plot_anomaly_decomposition(ncol = 3, alpha_dots = 0.7)
```
 time_decompose() functions:
 - #observed plot# : The overall observed data.
 - #season plot# : The seasonal or cyclic trend. 
 - #trend plot# : The long-term trend. The default is a span of 3 months.
 - #remainder plot# : It is used for analyzing the outliers.

 The red points indicate anomalies according to the anomalize function
 
 ##Recomposing##
 -The time_recompose() function is used to generate bands around the "normal" levels of observed values
```{r}
df %>%
time_decompose(Sales, method = 'stl', frequency = 'auto', trend = 'auto') %>%
anomalize(remainder, method = 'gesd', alpha = 0.1, max_anoms = 0.1) %>%
time_recompose() %>%
plot_anomalies(time_recomposed = TRUE, ncol = 3, alpha_dots = 0.5)
```

 - The plot and shows the anomalies
 
 - check actual anomalies values
```{r}
anomalies = df %>%
time_decompose(Sales, method = 'stl', frequency = 'auto', trend = 'auto') %>%
anomalize(remainder, method = 'gesd', alpha = 0.05, max_anoms = 0.1) %>%
time_recompose() %>%
filter(anomaly == 'Yes')

anomalies
```

The red values are the anomalies

##5. CONCLUSION##

Carrefour Kenya database records contain anomalies

##6. RECOMMENDATION##

The above anomalies should be invvestigated
